---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #4'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/01/2020"
output:
  pdf_document: default
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Load necessary libraries
library(caret)
library(data.table)
# Set master seed
set.seed(71554)

# Set filepaths for data ingestion
urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW4/data/"
fileTrain = "insurance_training_data.csv"
fileTest = "insurance-evaluation-data.csv"

# Read training file
DT <- fread(paste0(urlRemote, pathGithub, fileTrain),
            colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
                           'double', 'factor', 'double', rep('factor', 4L),
                           'integer', 'factor', 'double', 'integer',
                           rep('factor', 2L), 'double', 'integer', 'factor',
                           rep('integer', 2L), 'factor'))

# Convert character currencies to doubles
DT[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]

# Read and process the test/evaluation file
eval <- fread(paste0(urlRemote, pathGithub, fileTest),
              colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
                             'double', 'factor', 'double', rep('factor', 4L),
                             'integer', 'factor', 'double', 'integer',
                             rep('factor', 2L), 'double', 'integer', 'factor',
                             rep('integer', 2L), 'factor'))
eval[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]

# Number of training observations
ntrobs <- dim(DT)[[1]]
```

# Introduction
The assignment for HW4 is to analyze and model a dataset containing
approximately 8000 records representing customers of an auto insurance company.
Each record has two response variables. The first response variable,
`TARGET_FLAG`, is a 1 or a 0. where a `1` means that the person was in a car
and a `0` means that the person was not in a car crash. The second response
variable is `TARGET_AMT`. This value is 0 if the person did not crash their car.
However, if they did crash their car, this number will be a value greater than
zero.

The objective of the assignment is to build multiple linear regression and
binary logistic regression models on the training data to predict **both** the
probability that a person will crash their car and the amount of money it will
cost given a crash. Only the variables in the dataset, or variables directly
derived from them, may be used.

# DATA EXPLORATION
## Variables
The data is composed of the following variables:

|VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
|-|-|-|
|INDEX|Identification Variable (do not use)|None|
|TARGET_FLAG|Was Car in a crash? 1=YES 0=NO|None|
|TARGET_AMT|If car was in a crash, what was the cost|None|
|AGE|Age of Driver|Very young people tend to be risky. Maybe very old people also.|
|BLUEBOOK|Value of Vehicle|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
|CAR_AGE|Vehicle Age|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
|CAR_TYPE|Type of Car|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
|CAR_USE|Vehicle Use|Commercial vehicles are driven more, so might increase probability of collision|
|CLM_FREQ|# Claims (Past 5 Years)|The more claims you filed in the past, the more you are likely to file in the future|
|EDUCATION|Max Education Level|Unknown effect, but in theory more educated people tend to drive more safely|
|HOMEKIDS|# Children at Home|Unknown effect|
|HOME_VAL|Home Value|In theory, home owners tend to drive more responsibly|
|INCOME|Income|In theory, rich people tend to get into fewer crashes|
|JOB|Job Category|In theory, white collar jobs tend to be safer|
|KIDSDRIV|# Driving Children|When teenagers drive your car, you are more likely to get into crashes|
|MSTATUS|Marital Status|In theory, married people drive more safely|
|MVR_PTS|Motor Vehicle Record Points|If you get lots of traffic tickets, you tend to get into more crashes|
|OLDCLAIM|Total Claims (Past 5 Years)|If your total payout over the past five years was high, this suggests future payouts will be high|
|PARENT1|Single Parent|Unknown effect|
|RED_CAR|A Red Car|Urban legend says that red cars (especially red sports cars) are more risky. Is that true?|
|REVOKED|License Revoked (Past 7 Years)|If your license was revoked in the past 7 years, you probably are a more risky driver.|
|SEX|Gender|Urban legend says that women have less crashes then men. Is that true?|
|TIF|Time in Force|People who have been customers for a long time are usually more safe.|
|TRAVTIME|Distance to Work|Long drives to work usually suggest greater risk|
|URBANICITY|Home/Work Area|Unknown|
|YOJ|Years on Job|People who stay at a job for a long time are usually more safe|


