---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #3'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "10/27/2020"
output:
  pdf_document: default
urlcolor: purple
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(tidymodels)
library(data.table)
library(vip)
library(summarytools)
library(corrplot)
library(knitr)
library(rsample) # model 1 libraries
library(caret)
library(geoR)

set.seed(9450)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW3/data/"
fileTrain = "crime-training-data_modified.csv"
fileTest = "crime-evaluation-data_modified.csv"

df <- read.csv(paste0(urlRemote, pathGithub, fileTrain))
nobs <- dim(df)[[1]]
DT <- as.data.table(df)
DT[, target := factor(target)]
eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
```

# DATA EXPLORATION
## Data Description
The training data set contains 466 records summarizing attributes of various
neighborhoods in the city of Boston. The response variable is coded such that it
is `1` when the neighborhood’s crime rate is above the median and `0` when it is
not. In all, there are 12 predictors. These include:

Predictor Variables|Description   
-|----
zn | proportion of residential land zoned for large lots (over 25000 square feet) 
indus | proportion of non-retail business acres per suburb  
chas | a dummy var. for whether the suburb borders the Charles River (1) or not (0)
nox | nitrogen oxides concentration (parts per 10 million) 
rm | average number of rooms per dwelling   
age | proportion of owner-occupied units built prior to 1940   
dis | weighted mean of distances to five Boston employment centers  
rad | index of accessibility to radial highways   
tax | full-value property-tax rate per $10,000 
ptratio | pupil-teacher ratio by town 
lstat | lower status of the population (percent) 
medv | median value of owner-occupied homes in $1000s 

We can make some initial observaations. The data set has complete cases, thus, 
there is no need for imputation. Based on some common summary statistics, 
there are more observations where the crime rate is below the median. It is already 
apparent that some of the predictors varies depending the crime rate. For instance, 
there is a noticeable difference in the means of `age`, `lstat`, `rad`, and `zn` between 
the crime rate groups. 

```{r sumstat}
# DATA EXPLORATION 
summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
kable(summarystat[[1]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate > Median", 
      digit = 2L)
kable(summarystat[[2]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate < Median", 
      digit = 2L)
```

## Data Distribution
    
For each predictors, we computed and drew kernel density estimate to understand 
their distribution. Using density plots, we show how predictors are distributed 
between areas where crime rate is higher than the median, i.e. blue, and areas 
where crime rate is below the median, i.e. red. Our focus is to understand 
variables that highlight large variations between the two groups. 
  
```{r density, fig.width=8}
# density plots
DT[, IDX := .I]
DTM <- melt(DT, id.vars = c('IDX', 'target'), measure.vars = 1:12,
            variable.name = 'metric', value.name = 'value')
ggplot(DTM, aes(x = value, fill = target, color = target)) +
  geom_density(alpha = 0.4, show.legend = FALSE) +
  facet_wrap(~ metric, scales = 'free') 
```
  
The plots reveal that most of the data does not have a normal distribution. 
As expected, we see the heavily right tailed variables and others, which are 
multi-modal. The most Gaussian of the variables appears to be the one related 
to the average number of rooms per dwelling: `rm`. 
  
Another interesting predictor, `zn` i.e. proportion of residential land zoned 
for large lots, has a significant positive skew (skew = 2.18). Nearly 73% of 
the observations (339 of 466 total) have a value of zero. When analyzing the 
difference between the crime rate groups, it is possible that areas with high 
crime rate do not have land zoned for large lots suggesting suburban areas are 
likely to have low crime rates because there typically have large lots, whereas 
urban areas have a higher crime rate since lot size are smaller than 25000 square 
feet.  
  
Lastly, a third highly skewed variable is `chas`, the indicator as to whether 
the lot borders the Charles River. Out of the `r nobs` observations,
`r sum(DT$chas == 0)` do **not** border the river.  
  
To provide another view, this time highlighting outliers, we present the data and 
analyze using boxplots.
  
```{r boxplots, fig.width=8}
# boxplots 

ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, notch = TRUE,
               notchwidth = 0.8, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 3) + 
  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16,
               size = 2) + coord_flip() +
  labs(title = "Boxplot of Predictor Variables") + scale_y_log10()

ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(alpha = 0.5, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 2, aes(fill = target)) + 
  facet_wrap( ~ metric, scales = "free") + 
  labs(title = "Boxplot of Predictor Variables by Crime Rate") +
  scale_y_log10()
```

It is clear from the second boxplot, similar to the densities, that the
distribution of the predictor variables is different for for two outcomes. This
suggests that a model will be able to extract signal from the data.

## Data Correlation
  
```{r correlation, fig.height=5, fig.width=5}
# correlations
corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
```

Looking at the corrgram, we see that there are many variables which are
moderately to highly correlated, $\left|\rho\right| > 0.50$. Of note is that
`nox` has a largest positive correlation with `target` and `dis` has the largest
negative correlation with `target`. 
  
It could be possible we are seeing a reflection of the noticeable influence 
that socio-economic status has on the crime rate of an area.  Graif, Gladfelter 
and Matthews (2016) state crime is generally concentrated in disadvantaged, 
urban neighborhoods in the United States.  An economic segregation suggests 
that affluent neighborhoods may be further away in terms of distance, and as 
a result, disadvantaged areas are more attractive to crime because the probability 
of success is higher even if the targets are not as profitable. We prove this data, 
showing areas with a higher socio-economics correlating to a lower crime rate. 

Fields (2004) states industrial areas correlate with higher levels of nitrogen oxide 
concentrations. It may be an indication of why these areas are less dense with residents 
of a higher status. We see the same trend in `ptratio` since a higher ratio means less 
funding for public institutions, which is common in areas of lower status.  

([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4928692/](Source)).
([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1247398//](Source)).
  
Lastly, it seems that the variable `chas`, which indicate whether the suburb borders 
the Charles River, has statistically insignificant correlation with almost all of the 
other variables.

# DATA PREPARATION
  
As there are no missing values, we have no need to impute. Instead, different preparation
techniques are utilized for different models.

## Models 1 & 2
  
The variable `chas` will be removed due to having statistically insignificant correlation 
with almost all the variables in the dataset.  Next, outliers are tempered by capping below 
the 5th percentile and above the 95th percentile.  
  
```{r model1_2_adjust}
new_df <- subset(df, select = -chas)

# Cap all observations at their 5th and 95th percentiles
low5 <- apply(new_df, 2, quantile, prob = 0.05)
up95 <- apply(new_df, 2, quantile, prob = 0.95)
for (i in seq_along(new_df)) {
  new_df[, i] <- pmin(new_df[, i], up95[i])
  new_df[, i] <- pmax(new_df[, i], low5[i])
}
dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
dens_lstatO <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()
```

The Box-Cox transformation will be applied to the `age`, `lstat`, and `rm`
variables.

```{r m12BC}
ageBC <- boxcoxfit(new_df$age)
lstatBC <- boxcoxfit(new_df$lstat)
rmBC <- boxcoxfit(new_df$rm)
```

The transform suggests a \(\lambda\) of `r ageBC$lambda` for `age`,
`r lstatBC$lambda` for `lstat`, and `r rmBC$lambda` for `rm`

```{r m12applyBC}
new_df$age <- new_df$age ^ ageBC$lambda
new_df$lstat <- new_df$lstat ^ lstatBC$lambda
new_df$rm <- new_df$rm ^ rmBC$lambda
```

Below are the densities of the variables prior (left) and post (right) the
Box-Cox transform. The variables `lstat` and `rm` have benefited from the
transform, but the strong bimodality of `age` has not been affected
significantly.

```{r m12BC_plot}
dens_age <- ggplot(new_df, aes(age)) + geom_density()
dens_lstat <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rm <- ggplot(new_df, aes(rm)) + geom_density()
grid.arrange(dens_ageO, dens_lstatO, dens_rmO, dens_age, dens_lstat, dens_rm,
             layout_matrix = cbind(c(0, 1, 2), c(3, 4, 5)))
```

## Model 3
  
For model 3, we chose very little data manipulation, as there may be valuable 
information in the outliers.  A reasonable model may be able to extract 
something from `chas`.  Some variables, such as `chas` and `rad` (an index of 
accessibility) are converted to a factor. Additionally, `rad` probably ordered. 
Given there are now two factors in the predictors, we require creation of dummy 
variables. Lastly, the **labels** of `target` will be changed from integer to 
character for compliance with some of R's naming conventions.  Based on the 
definitions, 0 will be mapped to *below* and 1 to *above*.
  
# BUILD MODELS
  
First, we will split the training data into two sets. This will allow us to
perform a cross validation scheme on the models to tune parameters for optimum
performance. 

```{r m12_tts}
# BUILD MODELS
# train-test split

crime_df <- new_df %>%
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

# This is so Model 3 can use the same train/test split as Models 1 and 2
crime_df$id <- seq_len(nrow(crime_df))

# Changing the text above caused changes in output as fewer command were run. It
# is ALWAYS a good idea to set a seed IMMEDIATELY before any sampling in a
# Markdown file to ensure repeatability.

set.seed(7104)
data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df <- training(data_split)
test_df <- testing(data_split)

# Capture the indices and remove the no-longer-necessary column
tstIDX <- test_df$id
crime_df <- subset(crime_df, select = -id)
train_df <- subset(train_df, select = -id)
test_df <- subset(test_df, select = -id)
```

## Model 1
### Base Model
  
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset. 

```{r baseModel1}
baseModel <- glm(target~., family = binomial, data = train_df)
kable(summary(baseModel)$coefficients, digits = 3L,
      caption = 'Base Model Logistic Regression Output')
```

We can immediately see that a few variables *exceed* the 0.05 p-value threshold
for significance. 

### Enhanced Model
 
We will use **backwards stepwise regression** to remove features that are not
statistically significant in predicting the target. The result is a model that
includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`,
`ptratio`, `lstat`, and `medv`. 

### Coefficient Discussion
  
It is important to note that the coefficients from the model are predicting 
whether or not the target variable is a **1** (the crime rate *above* the 
median value). Additionally, since the numeric coefficients are relative to 
the range of values that the variable encompasses it is possible to have a 
coefficient that seems small when we look at the absolute magnitude, but that 
actually has a very strong effect when applied to the data.
  
```{r model1_coeff}
# 10-fold cross validation
train_control <- trainControl(method = "cv", number = 10)

# train the model on training set
model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat + medv,
               data = train_df,
               trControl = train_control,
               method = "glm",
               family=binomial())
kable(summary(model1)$coefficients, digits = 3L,
      caption = 'Backwards Model Regression Output')
```
  
Looking at the *signs* of the coefficients, we can see the variables `zn` 
and `tax` are negative. This is indicative of an inverse relationship; 
i.e., the higher these values, the less likely the crime rate is above the 
median. The relationship for the zone variable aligns with the findings from 
our initial data exploration. However, the relationship for the tax variable 
does not; in fact, we saw that there was a positive correlation between the 
two variables.  As the tax rate increases so does the probability that the 
crime rate will be above the median.  
  
The rest of the variables have positive coefficients and many of these are 
expected. For example, in our exploratory data analysis, we noticed that the 
`nox` variable (nitrogen oxides concentration (parts per 10 million)) has the 
greatest positive correlation with the target variable. This is surprising!  
We had not thought that the nitrogen oxide concentration would have as large 
of an impact on the crime rate. According to Colorado State University (2019), 
there may actually be a strong relationship between the two.  The team of 
researchers found strong correlation between exposure to air pollution and 
aggressive behavior.
  
([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).
  
## Model 2
### Logistic Classifier

```{r m2setup}
cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)

crime_recipe <- recipe(target ~., data=train_df)

crime_wf<- workflow() %>%
  add_recipe(crime_recipe)
```


```{r m2svm}
logit_specs <-
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

```

Tuning the models for best parameters. 

```{r m2stack}
logit_wf <-
  crime_wf %>%
  add_model(logit_specs)

crtl_grid<-control_grid(verbose = FALSE)

logit_results <-
  tune_grid(
    logit_wf,
    resamples = cv_folds,
    control = crtl_grid,
    grid = 10,
    save_pred = TRUE,
    save_workflow = FALSE
)
```


```{r m2predict}
best_model_auc <- select_best(logit_results, "roc_auc")
final_auc<-
  finalize_workflow(
    logit_wf,
    best_model_auc
)

final_model_pred<-
  final_auc %>%
  last_fit(data_split) %>%
  collect_predictions()

```

Computing the metrics for the logit model:

```{r m2Stats}
m2_roc<- yardstick::roc_auc(
  final_model_pred,
  truth = target,
  contains(".pred_1")
  )

m2_acc<- yardstick::accuracy(
  final_model_pred, 
  truth = target, 
  estimate = .pred_class
)

m2_recall<- yardstick::recall(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)

m2_precise<- yardstick::precision(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)
m2_metrics_df<-
  bind_rows(m2_roc, m2_acc, m2_recall, m2_precise)
```


```{r}
logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```
  
We again see that the primary predictor in the logistic regression is the `nox` 
with `rad` and `ptratio` rounding out the top three predictors staying in line 
with the base model. Living close to the center of town, having teachers, and 
not having polluted air are the largest predictors in terms of staying above median 
crime rates.  
  
## Model 3
```{r m3trainTemp}
DT_train <- DT[!(IDX %in% tstIDX)]
```
    
Model 3 will be a binary logistic regression allowing for all variables and 
selected pair-wise interactions pruned using both forward and backward stepwise 
regression based on AICc to select the optimal parameters. The captured indices 
of the test set for models 1 & 2 are used to split the data for model 3 as well. 
To address the possibility of over-fitting, we performed five-fold cross-validation 
on the training set.  With the vanilla logistic regression model in R, `glm` with 
a binomial family and a logit link, there are no other hyper parameters available 
for tuning.  More sophisticated logistic regression implementation have these 
parameters for fine-tuning.    
  
Considering *all* interactions would be foolhardy. Given the need for dummy variables, 
there will 19 predictors and \(_{19}C_2 = 171\) possible pairs, too much. Therefore, 
intentional selection of pairs is required.

Looking at the predictors, there are some, which logically may interact with each other, 
such as:   
* `indus`
	* The concentration of non-retail businesses in the area
* `nox`
    		* Nitrous Oxide is a common industrial pollutant. 

There is a high correlation in the training set between the two. Allowing for its 
interaction may temper the result of their relationship without losing more information.  
  
What may prove interesting is to look at some unsupervised clustering techniques to 
provide us with insights as to which variables may be "close" enough to warrant interactions. 
Instead of treating the observations as samples, we will treat the predictors as samples!
  
```{r m3Clust, fig.height=6}
predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
plot(predClust, xlab = "Raw Predictors")
```

It seems that there may be reason to consider an interaction between `chas` and
`nox`. Can that many industrial plants be on the river?   
  
Similarly with `indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
and pollution of industrial areas.  
  
Our selected starting model will include the following interactions:  
  
  * indus:nox
  * indus:lstat
  * chas:nox
  * ptratio:medv
  * rm:dis
    
### Train Model
  
```{r m3dataPrep}
DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad),
          target = factor(target, labels = c('Below', 'Above')))]
DT_test <- DT[IDX %in% tstIDX]
DT_train <- DT[!(IDX %in% tstIDX)]
```
```{r m3build}
trC <- trainControl(method = 'none', classProbs = TRUE,
                    summaryFunction = twoClassSummary)
dVars <- dummyVars(target ~ indus * nox +
                     indus * lstat +
                     chas * nox +
                     ptratio * medv +
                     rm * dis + . - IDX, data = DT_train, fullRank = TRUE)
DTtrnx <- predict(dVars, newdata = DT_train)
DTtrny <- DT_train$target
set.seed(1)
m3Fit <- train(x = DTtrnx, y = DTtrny, method = 'glmStepAIC', trControl = trC,
               family = 'binomial', trace = 0, direction = 'both')
```

### Coefficient Discussion
```{r m3Sum}
kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
             caption = 'Stepwise Interaction-Allowed Logistic Regression Output')
```
  
The selected model has some interesting results. First, the predictor with the 
greatest effect is `nox`.  Second, the only model variable which serves to 
**decrease** the crime rate from the baseline is `rm`. 

The interactions between `indus` and both `nox` and `lstat` are significant. 
As theorized, these interactions serve to temper their individual contributions to 
crime. Remember, a `1` indicates crime above the median, so positive coefficients 
indicate contributions to *increasing* the crime rate. A similar observation holds 
`rm`. While `dis` on its own did not decrease the model's AICc, its interaction 
to temper `rm` did help the model's deviance. No other selected interactions are 
left in the model.  
  
A number of variables fell out of the model, such as `chas`, `dis`, and `zn` 
among others. The remaining predictors are all significant, at least at the 10% 
level, except for `rad.24`. However, from the last trace of the stepping routine 
shown below, removing it would ***drastically*** reduce the models performance.  
  
Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as 
well. This indicates that both `rad = 2` and `rad = 6` should be considered 
indistinguishable from `rad = 1`.   

```
                    Df Deviance    AIC
<none>                   90.565 122.56
+ dis                1   89.191 123.19
+ zn                 1   89.558 123.56
+ tax                1   89.700 123.70
- ptratio            1   93.763 123.76
+ age                1   90.156 124.16
+ `nox:chasOnRiver`  1   90.224 124.22
- `rm:dis`           1   94.254 124.25
+ chas.OnRiver       1   90.273 124.27
+ rad.6              1   90.315 124.31
- lstat              1   94.364 124.36
+ `ptratio:medv`     1   90.494 124.49
+ rad.2              1   90.554 124.55
- rm                 1   95.001 125.00
- `indus:lstat`      1   95.633 125.63
- `indus:nox`        1   97.351 127.35
- indus              1   97.597 127.60
- medv               1   98.568 128.57
- rad.3              1  101.594 131.59
- rad.5              1  105.594 135.59
- rad.7              1  118.832 148.83
- nox                1  134.786 164.79
- rad.8              1  149.967 179.97
- rad.4              1  152.757 182.76
- rad.24             1  152.797 182.80
```

# SELECT MODELS
## Criteria
In binary classification, often, the measures of focus are precision and recall,
as opposed to accuracy. A measure combining this is the F1 score, defined as
twice the sum of precision and recall divided by their product. Another oft-used
metric is the area under the receiver operating curve, AUC. We will look at
the three metrics of accuracy, F1, and AUC, and select the model that performs
best in at least two of the three metrics.

## Performance 
Below we show the confusion matrices for the three models, followed by a table with
the selected performance. 

```{r comparison}
compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                        ACC = double(3),
                        F1 = double(3),
                        AUC = double(3))
```

### Model 1
```{r model1_perf}
# Model 1 Performance
model1_preds <- predict(model1, test_df, type = "raw")
model1_probs <- predict(model1, test_df, type = "prob")
colnames(model1_probs) <- c('pred0', 'pred1')

model1_results <- test_df %>%
  bind_cols(pred = model1_preds, model1_probs)

# Metrics
m1_roc <- yardstick::roc_auc(
  model1_results,
  truth = target,
  pred0 # select the prob class that corresponds to first level of target
)

m1_acc <- yardstick::accuracy(
  model1_results, 
  truth = target, 
  estimate = pred
)

m1_recall <- yardstick::recall(
  model1_results,
  truth = target, 
  estimate = pred
)

m1_precise <- yardstick::precision(
  model1_results,
  truth = target, 
  estimate = pred
)

m1_metrics_df<-
  bind_rows(m1_roc, m1_acc, m1_recall, m1_precise)

m1CM <- confusionMatrix(model1_preds, model1_results$target,
                        mode = 'prec_recall')
compTable[1, 2] <- m1_acc$.estimate
compTable[1, 3] <- 2 * m1_precise$.estimate * m1_recall$.estimate /
  (m1_precise$.estimate + m1_recall$.estimate)
compTable[1, 4] <- m1_roc$.estimate
m1CM$table
```

### Model 2

```{r m2TestSet}
final_model_pred %>%
  conf_mat(truth = target, estimate = .pred_class)

m2.metrics <- logit_wf %>%
  last_fit(data_split) %>%
  collect_metrics()

compTable[2, 2] <- m2.metrics[1, 3]
compTable[2, 3] <- 2 * m2_precise$.estimate * m2_recall$.estimate /
  (m2_precise$.estimate + m2_recall$.estimate)
compTable[2, 4] <- m2.metrics[2, 3]
```

### Model 3

```{r m3TestSet}
tstdVars <- dummyVars(target ~ indus * nox +
                        indus * lstat +
                        chas * nox +
                        ptratio * medv +
                        rm * dis + . - IDX, data = DT_test, fullRank = TRUE)
DTtstx <- predict(tstdVars, newdata = DT_test)
DTtsty <- DT_test$target
m3Pred <- predict(m3Fit, newdata = DTtstx)
m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')
m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
m3CM$table
compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
compTable[3, 3] <- m3CM$byClass[7]
compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
```

### Selected Performance Statistics
```{r selperfstat}
kable(compTable, digits = 3L, caption = 'Model Test Results')
```

All the models performed very well and classified the majority of the test set
properly. The difference between the three models was very slight. Only 7, 6, 
and 4 observations out of 92 were misclassified by the three models
respectively. Given the above statistics and confusion matrices, **Model 3**
barely edged out the other two and will be used to predict on the `eval` set.

## Predictions

```{r evalPred}
setDT(eval)
neval <- dim(eval)[[1]]
eval[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad))]
evalVars <- dummyVars( ~ indus * nox + indus * lstat + chas * nox + 
                         ptratio * medv + rm * dis + .,
                       data = eval, fullRank = TRUE)
evaltstx <- predict(evalVars, newdata = eval)
evalPred <- predict(m3Fit, newdata = evaltstx)
evalPredP <- predict(m3Fit, newdata = evaltstx, type = 'prob')
predTable <- data.frame(TestID = seq_len(neval),
                        PredictedClass = ifelse(evalPred == "Below", 0, 1),
                        Prob_Is_0 = evalPredP$Below,
                        Prob_Is_1 = evalPredP$Above)
kable(predTable, digits = 3L, caption = "Predicted Classes and Probabilities")
```


# APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.

```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r sumstat}
```
```{r density}
```
```{r boxplot}
```
```{r correlation}
```
```{r model1_2_adjust}
```
```{r m12BC}
```
```{r m12applyBC}
```
```{r m12BC_plot}
```
```{r model1_coeff}
```
```{r m2setup}
```
```{r m2svm}
```
```{r m2stack}
```
```{r m2predict}
```
```{r m2Stats}
```
```{r m3trainTemp}
```
```{r m3Clust}
```
```{r m3dataPrep}
```
```{r m3Sum}
```
```{r comparison}
```
```{r model1_perf}
```
```{r m2TestSet}
```
```{r m3TestSet}
```
```{r selperfstat}
```
```{r evalPred}
```